### 2021-04-30 Demographics Review R markdown.Rmd
Updates
Updates R markdown file with final scoring sheet. Edits to mean age plot and addition of funding plot.

# Download scoring sheet 
data.original <- read_excel("scoring_rhiannon.xlsx")
```

@@

# Compute percentage of articles that report funding number, out of those that report funding
percent.funding.num <- round(nrow(data.funding.num)/as.numeric(nrow(data.funding.name)) * 100, 2)
list.funding <- tolower(data.included$funding_names)

# Split the category combinations and convert to lower case letters
list.funding <- as.list(strsplit(list.funding, ","))
list.funding <- unlist(list.funding)

# Count number of occurrences for each category
table.list.funding <- as.data.frame(table(list.funding))
n.funding.nih <- sum(str_count(list.funding, "nih"), str_count(list.funding, "national institute"))

# Compute percentage of articles that report funding from the National Institutes in the US
data.funding.NI <- data.funding.name %>%
  filter(grepl("NI", funding_names) | grepl("National Institute", funding_names))
percent.funding.NI <- round(nrow(data.funding.NI)/as.numeric(nrow(data.included)) * 100, 2)

# Select articles that report funding number and funding from National Institutes in the US
data.funding.NI.num <- data.funding.name %>%
  filter(grepl("NI", funding_names) | grepl("National Institute", funding_names)) %>%
  filter(funding_refnum_yn == 1)
percent.funding.NI.num <- round(nrow(data.funding.NI.num)/as.numeric(nrow(data.funding.NI)) * 100, 2)
```

@@

# Heat map of reporting of age variables by year of publication
figure4A <- ggplot(data=age.corr.year, aes(x=variable, y=article_year, fill=value)) +
  geom_tile() +
  theme_apa() +
  ylab(label="Year") +
  xlab("Age variable") +
  scale_fill_viridis_c(end=0.6, direction=-1, labels=c("0.0","0.25","0.5","0.75","1")) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8), 
        legend.position="none", legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,-10,-10), legend.title=element_blank())

@@

# Heat map of reporting of age variables by journal of publication
figure4B <- ggplot(data=age.corr.journal, 
                   aes(x=variable, y=article_journal, fill=value)) +
  geom_tile() +
  theme_apa() +
  ylab(label="Journal") +
  xlab("Age variable") +
  scale_fill_viridis_c(end=0.6, direction=-1, labels=c("0.0","0.25","0.5","0.75","1")) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.position="right", legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,-10,-10), legend.title=element_blank())

@@

# Change age from weeks to years, with 1 month = 4.3 weeks, and from months to years
# age not in years is entered with "months" or "weeks"
data.age.mean.not.years <- data.age.mean.not.years %>%
  separate(age_reported_mean_value, 
           into=c("age_reported_mean_value", "text"),
           sep=" ")
data.age.mean.not.years$age_reported_mean_value <- as.numeric(data.age.mean.not.years$age_reported_mean_value)

data.age.mean.not.years <- data.age.mean.not.years %>%
  mutate(age_reported_mean_value=ifelse(text=="weeks", 
                                        age_reported_mean_value / 4.3 /12,
                                        age_reported_mean_value / 12)) %>%
  select(article_pmid, article_year, age_reported_mean_value)# else case is in months

@@

# Violin plot with regression line and equation for mean age by year of publication
figure4C <- ggplot(data=data.age.mean, aes(x=article_year, y=age_reported_mean_value)) +
  geom_smooth(method="lm", color="#440154FF") +
  geom_point(size=1, color="#33638DFF") +
  geom_text(x=1988, y=10, size=3.3,
            label="Mean age = 281.5 - 0.1 * Year of publication") + # edit manually
  geom_text(x=1988, y=7, size=3.1, label="R-Sq = 0.3%") + # edit manually
  theme_apa() +
  ylab("Mean age (years)") +
  xlab("Year") +
  scale_x_continuous(limits=c(1979,2019),breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(limits=c(0,60), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8))
```

@@

# Stacked bar chart  of reporting of other demographic variables by year of publication
figure6.bars <- ggplot(data.demo, aes(x=variable, y=per, fill=report)) +
  geom_bar(stat="identity") +
  facet_grid(~article_year) +
  theme_apa() +
  xlab("Year of publication") +
  ylab("Proportion of articles") +
  labs(fill="Variable reported") +
  scale_fill_manual(values=c("#482677FF","#D8DE29FF")) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"),
        axis.text.x=element_text(size=5, angle=45, hjust=1),
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.title=element_text(face="bold", size=12),
        legend.margin=margin(0,0,0,0), legend.box.margin=margin(-10,0,-10,-10),
        panel.spacing=unit(0.1, "lines"))

data.demo.reported <- data.demo %>%
  filter(report == "No") %>%
  mutate(per = (100 - per))

figure6.points <- ggplot(data.demo.reported, aes(x=article_year, y=per, color=factor(variable))) +
  geom_point(size=2.5) +
  geom_line(linetype="dashed") +
  theme_apa() +
  xlab("Year of publication") +
  ylab("Proportion of articles") +
  scale_color_viridis_d(name="Variable", 
                        labels=c("Education","Ethnicity","Profession","Socio-economic status")) +
  scale_x_continuous(limits=c(1979,2019), breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(labels=c("0%","25%","50%","75%","100%"), limits=c(0,100)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"),
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.title=element_text(face="bold", size=12),
        legend.margin=margin(0,0,0,0), legend.box.margin=margin(-10,0,-10,-10))

@@

# Compute distribution of the number of categories reported for education
data.education$education_reported_numcategory <- as.numeric(data.education$education_reported_numcategory)
figure7A <- ggplot(data=data.education, aes(x=education_reported_numcategory)) +
  geom_histogram(binwidth=1, aes(y=..density..), colour="black", fill="#2D708EFF") +
  geom_vline(aes(xintercept=mean(education_reported_numcategory, na.rm=TRUE)), 
             color="#D8DE29FF", size=1) +
  theme_apa() +
  ylab("Density") +
  xlab("Education") +
  scale_x_continuous(limits=c(0.5,8.5), breaks=c(1,2,3,4,5,6,7,8)) +
  scale_y_continuous(limits=c(0,0.5), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"),
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8))

@@

  Funding sources were reported by `r round(percent.funding.name)`% of studies, while funding number was also reported in `r round(percent.funding.num)`% of these cases. Overall, funding by the United States' National Institutes of Health (NIH) represented `r round(percent.funding.NI)`% of the reported funding agencies. `r round(percent.funding.NI.num)`% of the studies funded by the NIH also reported funding number.

@@

```{r, fig.cap="Geographical location of the studies. The eight most represented countries are individually shown.", warning=FALSE, message=FALSE}
figure2B
```

@@

```{r, fig.cap="Reporting of education, ethinicity, profession and socio-economic status.", warning=FALSE, message=FALSE}
#figure6.bars
figure6.points
```

>>>>

# Download scoring sheet 
data.original <- read_excel("scoring.xlsx")
```

@@

# Compute percentage of articles that report funding number, out of those that report funding
percent.funding.num <- round(nrow(data.funding.num)/as.numeric(nrow(data.funding.name)) * 100, 2)

# Compute whether articles report funding and funding number
data.funding <- data.included %>%
  mutate(funding=if_else(!is.na(funding_names) & funding_refnum_yn==0, 
                         "Funding Reported", 
                         if_else(!is.na(funding_names) & funding_refnum_yn==1, 
                                 "Funding and Grant Reported", 
                                 "Funding Not Reported")))

data.funding$funding <- as.factor(data.funding$funding)

# Count number of articles for funding condition by year of publication
data.funding <- data.funding %>%
  count(funding, article_year) %>%
  ungroup()

# Compute proportion of articles per sex category by year of publication
data.funding <- data.funding %>%
  group_by(article_year) %>%
  mutate(total=sum(n)) %>%
  group_by(funding, add=TRUE) %>%
  mutate(per=round(100*n/total,3))
data.funding$article_year <- as.numeric(data.funding$article_year)
data.funding$per <- as.numeric(data.funding$per)

# Stacked bar chart of reporting of funding and funding number by year of publication
figure2A <- ggplot(data.funding, aes(x=article_year, y=per)) +
  geom_col(aes(fill=factor(funding, levels=c("Funding Not Reported",
                                            "Funding Reported",
                                            "Funding and Grant Reported")))) +
  theme_apa() +
  xlab("Year of publication") +
  ylab("Proportion of articles") +
  labs(fill="") +
  scale_fill_manual(values=c("#482677FF","#DCE319FF","#FDE725FF")) +
  scale_x_continuous(breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(labels=c("0%","25%","50%","75%","100%"), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.margin=margin(0,0,0,0), legend.box.margin=margin(-10,0,-10,-10))

# Change funding names to lower case
data.funding.name$funding_names <- tolower(data.funding.name$funding_names)

# Create list of funding sources
list.funding <- data.funding.name$funding_names

# Split combinations of funding sources (multiple funding names)
# which are delimited by commas
list.funding <- as.list(strsplit(list.funding, ","))
list.funding <- unlist(list.funding)

# Count number of occurrences for each funding name
table.list.funding <- as.data.frame(table(list.funding))

# Combine funding names that fall under the same agency
table.list.funding$list.funding <- as.character(table.list.funding$list.funding)

# NIH (USA)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("nih", table.list.funding$list.funding)== TRUE | grepl("national institute", table.list.funding$list.funding) == TRUE | grepl("nimh", table.list.funding$list.funding)== TRUE | grepl("nhlbi", table.list.funding$list.funding)== TRUE | grepl("national heart lung", table.list.funding$list.funding)== TRUE | grepl("national cancer institute", table.list.funding$list.funding)== TRUE | grepl("nia", table.list.funding$list.funding)== TRUE | grepl("ninds", table.list.funding$list.funding)== TRUE | grepl("ninr", table.list.funding$list.funding)== TRUE | grepl("ninda", table.list.funding$list.funding)== TRUE, 
                              "NI", table.list.funding$list.funding))
# NHMRC (AUS)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("nhmrc", table.list.funding$list.funding)== TRUE | grepl("national health and medical research council", table.list.funding$list.funding)== TRUE,
                              "NHMRC", table.list.funding$list.funding))

# Wellcome Trust (GBR)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("wellcome", table.list.funding$list.funding)== TRUE,
                              "Wellcome Trust", table.list.funding$list.funding))

# CIHR (CAN)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("canadian institute", table.list.funding$list.funding)== TRUE,
                              "CIHR", table.list.funding$list.funding))

# NSFC (CHN)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("national natural science", table.list.funding$list.funding)== TRUE,
                              "NSFC", table.list.funding$list.funding))

# Academy of Finland (FIN)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("academy of finland", table.list.funding$list.funding)== TRUE,
                              "Academy of Finland", table.list.funding$list.funding))

# Inserm (FRA)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("inserm", table.list.funding$list.funding)== TRUE,
                              "Inserm", table.list.funding$list.funding))

# SNSF (CHE)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("swiss national", table.list.funding$list.funding)== TRUE,
                              "SNSF", table.list.funding$list.funding))

# Department of Veterans Affairs (USA)
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("veterans affairs", table.list.funding$list.funding)== TRUE,
                              "Dept Vet Affairs", table.list.funding$list.funding))

# European Commission
table.list.funding <- table.list.funding %>%
  mutate(list.funding=if_else(grepl("european commission", table.list.funding$list.funding)== TRUE,
                              "EC", table.list.funding$list.funding))

# Group by funding agency
table.list.funding <- table.list.funding %>%
  group_by(list.funding) %>%
  summarise(N = sum(Freq))

# Compute percentage of articles that report funding from the National Institutes in the US
data.funding.NI <- data.funding.name %>%
  filter(grepl("nih", funding_names) | grepl("national institute", funding_names) | grepl("nimh", funding_names) | grepl("nhlbi", funding_names) | grepl("national heart lung", funding_names) | grepl("national cancer institute", funding_names) | grepl("nia", funding_names) | grepl("ninr", funding_names) | grepl("ninda", funding_names) | grepl("ninds", funding_names))

percent.funding.NI <- round(nrow(data.funding.NI)/as.numeric(nrow(data.included)) * 100, 2)

# Select articles that report funding number and funding from National Institutes in the US
data.funding.NI.num <- data.funding.NI %>%
  filter(funding_refnum_yn == 1)

percent.funding.NI.num <- round(nrow(data.funding.NI.num)/as.numeric(nrow(data.funding.NI)) * 100, 2)
```

@@

# Heat map of reporting of age variables by year of publication
figure4A <- ggplot(data=age.corr.year, aes(x=variable, y=article_year, fill=value)) +
  geom_tile() +
  theme_apa() +
  ylab(label="Year") +
  xlab("Age variable") +
  scale_fill_viridis_c(end=0.6, direction=-1, name="Reporting",
                       limits=c(-1,1), labels=c("0.0","0.25","0.5","0.75","1")) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8), 
        legend.position="right", legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,-10,-10), legend.title=element_text(face="bold"))

@@

# Heat map of reporting of age variables by journal of publication
figure4B <- ggplot(data=age.corr.journal, 
                   aes(x=variable, y=article_journal, fill=value)) +
  geom_tile() +
  theme_apa() +
  ylab(label="Journal") +
  xlab("Age variable") +
  scale_fill_viridis_c(end=0.6, direction=-1) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.position="none", legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,-10,-10), legend.title=element_blank())

@@

# Change age from days / weeks / months to years, with 1 month = 4.3 weeks = 30.4 days
# age not in years is entered with "months" or "weeks"
data.age.mean.not.years <- data.age.mean.not.years %>%
  separate(age_reported_mean_value, 
           into=c("age_reported_mean_value", "text"),
           sep=" ")
data.age.mean.not.years$age_reported_mean_value <- as.numeric(data.age.mean.not.years$age_reported_mean_value)

data.age.mean.not.years <- data.age.mean.not.years %>%
  mutate(age_reported_mean_value=ifelse(text=="weeks", 
                                        age_reported_mean_value / 4.3 / 12,
                                        if_else(text=="days",
                                                age_reported_mean_value / 30.4 / 12,
                                                age_reported_mean_value / 12))) %>%
  select(article_pmid, article_year, age_reported_mean_value)# else case is in months

@@

# Violin plot with regression line and equation for mean age by year of publication
figure4C <- ggplot(data=data.age.mean, aes(x=article_year, y=age_reported_mean_value)) +
  geom_smooth(method="lm", color="#440154FF") +
  geom_point(size=1, color="#33638DFF") +
  #geom_text(x=1988, y=10, size=3.3,
            #label="Mean age = 281.5 - 0.1 * Year of publication") + # edit manually
  #geom_text(x=1988, y=7, size=3.1, label="R-Sq = 0.3%") + # edit manually
  theme_apa() +
  ylab("Mean age (years)") +
  xlab("Year of publication") +
  scale_x_continuous(limits=c(1979,2019),breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(limits=c(0,80), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"), 
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8))
```

@@

# Scatter plot of reporting of other demographic variables by year of publication
data.demo.reported <- data.demo %>%
  filter(report == "No") %>%
  mutate(per = (100 - per))

figure6 <- ggplot(data.demo.reported, aes(x=article_year, y=per, color=factor(variable))) +
  geom_point(size=2.5) +
  geom_line(linetype="dashed") +
  theme_apa() +
  xlab("Year of publication") +
  ylab("Proportion of articles") +
  scale_color_viridis_d(name="Variable", 
                        labels=c("Education","Ethnicity","Profession","Socio-economic status")) +
  scale_x_continuous(limits=c(1979,2019), breaks=c(1979,1989,1999,2009,2019)) +
  scale_y_continuous(labels=c("0%","25%","50%","75%","100%"), limits=c(0,100)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"),
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8),
        legend.title=element_text(face="bold", size=12),
        legend.margin=margin(0,0,0,0), legend.box.margin=margin(-10,0,-10,-10))

@@

# Compute distribution of the number of categories reported for education
data.education$education_reported_numcategory <- as.numeric(data.education$education_reported_numcategory)

figure7A <- ggplot(data=data.education, aes(x=education_reported_numcategory)) +
  geom_histogram(binwidth=1, aes(y=..density..), colour="black", fill="#2D708EFF") +
  geom_vline(aes(xintercept=mean(education_reported_numcategory, na.rm=TRUE)), 
             color="#D8DE29FF", size=1) +
  theme_apa() +
  ylab("Density") +
  xlab("Education") +
  scale_x_continuous(limits=c(0.5,8.5), breaks=c(1,2,3,4,5,6,7,8)) +
  scale_y_continuous(limits=c(0,0.5), expand=c(0,0)) +
  theme(axis.title=element_text(face="bold", size=12), 
        axis.text=element_text(colour="black", face="bold"),
        axis.line=element_line(size=0.8), axis.ticks=element_line(size=0.8))

@@

  Funding sources were reported by `r round(percent.funding.name)`% of studies, while funding number was also reported in `r round(percent.funding.num)`% of these cases. Overall, funding by the United States' National Institutes of Health (NIH) represented `r round(percent.funding.NI)`% of the reported funding agencies. `r round(percent.funding.NI.num)`% of the studies funded by the NIH also reported funding number. The second most represented funding agencies were the Australian National Health and Medical Research Council (NHMRC) and the Canadian Institutes of Health Research (CIHR).

```{r}
figure2A
```

@@

```{r, fig.cap="Funding sources and geographical location of the studies. A. Reporting of funding across the years. B. Distribution of articles by study location. The eight most represented countries are individually shown.", warning=FALSE, message=FALSE}
figure2B
```

@@

```{r, fig.cap="Reporting of education, ethinicity, profession and socio-economic status.", warning=FALSE, message=FALSE}
figure6
```
